### 2023-08-29 manuscript.tex
Moved title page information to manuscript and added required sections for submission

\begin{abstract}
\noindent
Identifying protein sequences with common ancestry is a core task in bioinformatics and evolutionary biology. However, methods for inferring and aligning such sequences in annotated genomes have not kept pace with the increasing scale and complexity of the available data. Thus, in this work we implemented several improvements to the traditional methodology that more fully leverage the redundancy of closely related genomes and the organization of their annotations. Two highlights include the application of the more flexible \textit{k}-clique percolation algorithm for identifying clusters of orthologous proteins and the development of a novel technique for removing poorly supported regions of alignments with a phylogenetic HMM. In making the latter, we wrote a fully documented Python package Homomorph that implements standard HMM algorithms and created a set of tutorials to promote its use by a wide audience. We applied the resulting pipeline to a set of 33 annotated \textit{Drosophila} genomes, generating 22,813 orthologous groups and 8,566 high-quality alignments.
\end{abstract}

@@

% To investigate these hypotheses, the genome sequences and supporting transcript evidence for several segments and regions were examined in detail: 0254, 26CF, 202C, 188E, 2E9F, 34B7, 0C31.

@@

\subsection{Code and data availability}
\begin{sloppypar}
The code used to produce the results and analyses is available at \url{https://github.com/marcsingleton/orthology_inference2023}. HMM algorithms were implemented in the standalone package Homomorph which is available at \url{https://github.com/marcsingleton/homomorph} and on the Python Package Index (PyPI). The following Python libraries were used: matplotlib~\cite{Hunter2007}, NumPy~\cite{Harris2020}, pandas~\cite{McKinney2010}, and SciPy~\cite{Virtanen2020}. There are no primary data associated with this manuscript. All primary data are available from publicly accessible sources described in their corresponding sections. Relevant outputs are available in the supplemental files at FigShare. File S1 contains the orthologous groups generated by 4-clique percolation with paralogs added. File S2 contains the initial alignments of all sequences in each orthologous group. File S3 contains the refined alignments of the representative sequences in each single copy orthologous group. File S4 contains the insertion phylo-HMM model parameters and the alignments of the single copy ortholog groups after trimming. File S5 contains the missing data phylo-HMM model parameters and the missing segments identified for each sequence in each single copy orthologous group. Missing segments are given as Python slices in the following form: start0-stop0,start1-stop1,\ldots,startn-stopn. If the slices field is empty, then no missing segments were identified in that sequence. File S6 contains the consensus phylogenetic trees fit using the LG, GTR, and GTR2 substitution models.
\end{sloppypar}

>>>>

\textbf{Title:} Leveraging genomic redundancy to improve inference and alignment of orthologous proteins

\textbf{Authors:} Marc Singleton\textsuperscript{1,*}, Michael Eisen\textsuperscript{1,2}

\textbf{Author addresses:}\\
1 Howard Hughes Medical Institute, UC Berkeley, Berkeley, CA 94720, USA\\
2 Department of Molecular and Cell Biology, UC Berkeley, Berkeley, CA 94720, USA

\textbf{Corresponding author email address:}\\
* marcsingleton@berkeley.edu

\textbf{Running head:} Leveraging genomic redundancy

\textbf{Keywords:} orthology, multiple sequence alignment, HMM, Drosophila

\clearpage

\section{Abstract}
Identifying protein sequences with common ancestry is a core task in bioinformatics and evolutionary biology. However, methods for inferring and aligning such sequences in annotated genomes have not kept pace with the increasing scale and complexity of the available data. Thus, in this work we implemented several improvements to the traditional methodology that more fully leverage the redundancy of closely related genomes and the organization of their annotations. Two highlights include the application of the more flexible \textit{k}-clique percolation algorithm for identifying clusters of orthologous proteins and the development of a novel technique for removing poorly supported regions of alignments with a phylogenetic HMM. In making the latter, we wrote a fully documented Python package Homomorph that implements standard HMM algorithms and created a set of tutorials to promote its use by a wide audience. We applied the resulting pipeline to a set of 33 annotated \textit{Drosophila} genomes, generating 22,813 orthologous groups and 8,566 high-quality alignments.

\section{Article summary}
insert article summary

@@

@@

\section{Data availability}
\begin{sloppypar}
The code used to produce the results and analyses is available at \url{https://github.com/marcsingleton/orthology_inference2023}. HMM algorithms were implemented in the standalone package Homomorph which is available at \url{https://github.com/marcsingleton/homomorph} and on the Python Package Index (PyPI). The following Python libraries were used: matplotlib~\cite{Hunter2007}, NumPy~\cite{Harris2020}, pandas~\cite{McKinney2010}, and SciPy~\cite{Virtanen2020}. There are no primary data associated with this manuscript. All primary data are available from publicly accessible sources described in their corresponding sections. Relevant outputs are available in the supplemental files at FigShare. File S1 contains the orthologous groups generated by 4-clique percolation with paralogs added. File S2 contains the initial alignments of all sequences in each orthologous group. File S3 contains the refined alignments of the representative sequences in each single copy orthologous group. File S4 contains the insertion phylo-HMM model parameters and the alignments of the single copy ortholog groups after trimming. File S5 contains the missing data phylo-HMM model parameters and the missing segments identified for each sequence in each single copy orthologous group. Missing segments are given as Python slices in the following form: start0-stop0,start1-stop1,\ldots,startn-stopn. If the slices field is empty, then no missing segments were identified in that sequence. File S6 contains the consensus phylogenetic trees fit using the LG, GTR, and GTR2 substitution models.
\end{sloppypar}

\section{Acknowledgements}
insert acknowledgements

\section{Funding}
insert funding

\section{Conflicts of interest}
The authors declare no conflicts of interest.

\section{Figures}
insert figures

\section{Tables}
insert tables
